### 2022-07-06 supplemental/OJEMB Supplementary Material.tex
Add to supplemental info

%% Supplementary Materials IEEE OJEMB

\documentclass[journal]{IEEEtran}
\usepackage[usenames]{color}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{cite}
\usepackage{array}
\usepackage{pslatex} 
\usepackage{url}
\usepackage{lineno}
\usepackage{graphicx}  % Written by David Carlisle and Sebastian Rahtz
\usepackage{setspace}
\usepackage{tikz}
\usepackage{hhline}
\usepackage{mathtools}
\usepackage{caption}
\usepackage[letterpaper]{geometry}
\geometry{verbose,tmargin=0.7in,bmargin=0.7in,lmargin=0.65in,rmargin=0.65in}
\setlength{\headheight}{17pt}
\setlength{\headsep}{5pt}
%\captionsetup{labelfont={up},font=small}
\captionsetup[figure]{name={Fig.},labelsep=period,font=small}
 \captionsetup[table]{name={TABLE},labelsep=period,font=small}
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

\usepackage{lipsum}

\usepackage{soul}
\newcommand{\etal}{\textit{et al}., }
\newcommand{\ie}{\textit{i}.\textit{e}., }
\newcommand{\eg}{\textit{e}.\textit{g}.\ }
\newcommand{\cf}{\textit{cf}.\ }
\usepackage{hyperref}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

\newcommand{\flogo}{\includegraphics[height=18pt]{embLogo.eps}}

%\lhead{\textsc{\flogo}}

% The paper headers
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{\textsc{\flogo}}
\rhead{\textcolor{violet}{\Large\textbf{Supplementary Materials}}}
\fancyfoot[C]{\thepage}
\renewcommand{\footrulewidth}{0.4pt}% default is 0pt

\begin{document}

\twocolumn[
\begin{@twocolumnfalse}
\thispagestyle{fancy}   
\vspace*{0.5cm}
\textbf{\LARGE{Supplementary Materials}}\\
\\
\medskip\Large{Compressed Sensing for Characterization of Tinnitus}\\
\normalsize{ Alec Hoyland,
	Nelson Barnett, Benjamin W. Roop, Adam C. Lammert}\\
	\vspace*{0.5cm}
\end{@twocolumnfalse}
]

% make the title area

\IEEEPARstart{T}his document contains a description of additional experiments to find a performant
stimulus generation method and a discussion of sparsity in tinnitus signals.
The code for all experiments is freely available at \url{https://github.com/alec-hoyland/tinnitus-project}
and the data are available upon request.

\section{Stimulus Generation}

In the context of this paper, a stimulus generation method
is a process that generates a random waveform that is:

\begin{enumerate}
  \item auditorally-distinguishable
  \item statistically uncorrelated, and
  \item similar to tinnitus percepts.
\end{enumerate}

Additionally, compressed sensing requires that the matrix of stimuli
should satisfy the restricted isometry property,
which many random matrices do with high probability (\eg Gaussian random matrices)
\cite{candesRestrictedIsometryProperty2008,candesIntroductionCompressiveSampling2008}.

\subsection{Auditorally-Distinguishable Stimuli}

We used mel-frequency binning to ensure that our stimuli were auditorally-distinguishable.
The mel scale is a perceptual scale of pitches judged by listeners to be equal in distance
from one to another (Fig. \ref{fig:hz2mel}) \cite{hamiltonTopographySpeechrelatedAcoustic2020}.

The formula

\begin{equation}
  m = 2595 \log_10 \bigg( 1 + \frac{f}{700} \bigg)
\end{equation}

converts $f$ Hz to $m$ mels.

\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{gfx/hz2mel.eps}
	\caption{The relationship between Hz and mels is logarithmic.}
	\label{fig:hz2mel}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{gfx/example_stimulus.eps}
	\caption{Example stimulus. \textbf{(a)} shows the waveform of the stimulus,
  \textbf{(b)} the frequency spectrum from which the waveform was generated,
  and \textbf{(c)} the $100$-dimensional bin representation.}
	\label{fig:examplestimulus}
\end{figure}

Furthermore, to reduce the system complexity by more than 80x,
we implement tonotopic binning, where the frequency scale is binned
along $100$ equally mel-spaced bins (Fig. \ref{fig:examplestimulus}).

\section{Hyperparameter Sweep}

To determine a suitable stimulus generation method,
we performed a hyperparameter sweep over hyperparameters
of nine different stimulus generation methods.
We used an \textit{in-silico} model of the experiment,
matching against an American Tinnitus Association tinnitus example,
as in the human experiment.

The methods are as follows:

\begin{enumerate}
  \item \textit{Bernoulli}: A binned method in which each tonotopic bin
  has a probability $p$ of being set to $0$ dB otherwise it is set to $-20$ dB.
  \item \textit{Brimijoin}: A binned method inspired by \cite{brimijoinInternalRepresentationVowel2013}.
  Each bin is assigned an amplitude value chosen from a uniform distribution of discrete values
  on the interval $[-20, 0]$. We used $6$ steps, which is consistent with the original paper.
  \item \textit{Gaussian Noise No Bins}: The amplitude of each frequency was determined by
  Gaussian noise with known mean and variance, which are hyperparameters $\mu$ and $\sigma$.
  \item \textit{Gaussian Noise}: A binned method where the bin amplitudes were
  determined by a Gaussian random variable with known mean and variance, \eg $\mathcal{N}(\mu, \sigma)$.
  \item \textit{Gaussian Prior}: A binned method where the number of filled bins
  was set by $\mathrm{round}(\mathcal{N}(\mu, \sigma))$,
  and that many bins were filled randomly at $0$ dB (unfilled bins were set to $-20$ dB).
  \item \textit{Power Distribution}: A binned method where the amplitude of each bin
  is drawn from a distribution matching the histogram of amplitudes
  of the American Tinnitus Association tinnitus examples from their website.
  \item \textit{Uniform Noise No Bins}: The amplitude associated with each frequency
  is determined by a uniform random variable on the interval $[-20, 0]$ dB.
  \item \textit{Uniform Noise}: The amplitude associated with each tonotopic bin
  is determined by a uniform random variable on the interval $[-20, 0]$ dB.
  \item \textit{Uniform Prior}: The number of filled bins is determined by drawing an integer
  from the discrete interval $[x_1, x_2] \in \mathbb{Z}^+$. Then, that many bins are set to $0$ dB,
  and all other bins are set to $-20$ dB. $x_1$ and $x_2$ are hyperparameters.
\end{enumerate}

9 methods and 8 hyperparameters were varied in the hyperparameter sweep grid search.
Methods were evaluated \textit{in-silico} using the model:

\begin{equation}
  y = \mathrm{sgn}(\Psi * \bar{x})
\end{equation}

where $y \in \mathbb{R}^n$ is the binary response vector,
$\Psi \in \mathbb{R}^{n \times d}$ is the stimulus matrix,
and $\bar{x} \in \mathbb{R}^d$ is the true target signal (the spectrum of the ATA tinnitus example).

Each stimulus generation method, with different hyperparameters,
was evaluated for target signals: buzzing, electric, roaring, screeching, static, and tea kettle.

\begin{table}[ht]
	\begin{center}	
		\small
		\resizebox{1\columnwidth}{!}{
			\begin{tabular}{ m{0.35\columnwidth} m{0.65\columnwidth} }	
				\multicolumn{2}{c}{\textsc{Units for Magnetic Properties}} \\
				\hhline{===}
				{Symbol} & {Quantity} \\
				\hhline{===}

			\end{tabular}
		}
	\end{center}
\end{table}

% n_bins: [10 30 100 200 300]
% bin_prob: [0.1000 0.3000 0.5000 0.8000]
% amplitude_mean: -35
% amplitude_var: [5 10 20]
% n_bins_filled_mean: [1 3 10 20 30]
% n_bins_filled_var: [0.0100 1 3 10]
% min_bins: [1 3 10 20 30]
% max_bins: [10 20 30 50]

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,../paper/mybib}

% that's all folks
\end{document}